I owe a great deal to everyone who has influenced and supported me---without you this thesis wouldn't exist.  My parents, in particular, have always served as an example of who I want to be.  Thank you for bringing me to this country, for always encouraging me, and for your overall overachievement that serves as an ever-moving goal to shoot for.  The rest of my family have also been good examples: my grandparents embodying quiet intelligence, iron will, artistic sensibility, and pure kindness; my aunt and uncle displaying a great talent for \'{e}lan. Thank you for raising and teaching me.  To my cousins, who are as close as siblings: than you for being a constant source of surprise and pride.   Many thanks to all of my friends, who ensure that life is never dull.

I have also had the rare privilege of working with two of the finest researchers that I have ever encountered.  Jim Bell and Jeremy Kasdin both share a clear love and enthusiasm for their chosen fields and a drive to expand and share their knowledge.  I cannot thank you enough for your guidance and support.  

Those portions of this text derived from published papers benefited greatly from the input of my various co-authors, editors, and collaborators; any remaining mistakes are solely my own.  Jeremy Kasdin deserves a medal for reading every draft of every paper I have written in the last five years.  Bob Vanderbei's  comments and suggestions have vastly improved several of my papers and Eric Cady's erudite knowledge of arcane integrals allowed us to find some of the most interesting results presented here.  Stephanie Goldfarb's basic algorithm for generating reversion coefficients allowed me to write the code used to prove those results.  Tyler Groff deserves many thanks for patiently explaining optics to me (over and over again), and not over-using a certain file folder.  My readers, Ed Turner and Michael Littman, provided invaluable input into both this thesis and my research in general.

Many thanks go to Jessica O'Leary, who made navigating graduate school infinitely easier, and who doesn't spend nearly as much time on Facebook as I tell people.  Since her departure, Candy Reed and Jill Ray have done an excellent job in assuming both her official duties, and maintaining the community feeling that makes the MAE department such a great place to work.  I also owe much to Donald Knuth, Leslie Lamport, and the entire online \LaTeX community, who have freed me from the tyranny of Word forever.

The only remaining thanks are the most difficult, as they go to the one person whose contribution is incalculable.  Melanie, I dedicate this thesis to you.  Without you, life would be nowhere near as sweet.

\vfill
\noindent This dissertation carries the number T-3230 in the records of the Department of Mechanical and Aerospace Engineering.